---
title: "epigenetics_final"
author: "Brennan Baker"
date: "April 1, 2019"
output: html_document
---


```{r}
library(tidyverse)
library(broom)
library(corrplot)
library(psych)
```

Data import.

```{r}
df = read_csv(file = "C:/Users/brenn/Desktop/PHD 1/Epigenetics/epigenetics_final/data/Simulated_miRNA_Dataset_EHSC_P8326_Feb2019_FULL.csv")
```

For all the miRs we can use sapply, which applies a function to an entire dataset and provides the names of each column
```{r}
NAS = sapply(df, function(x) sum(is.na(x)))
```

Remove the miRs with greater than 70% NA
First select only the columns with greater than 70%
```{r}
NAS.70 = NAS[NAS/1000 >= 0.7]
#Then filter the dataset by the columns in that new vector
#"!" operator means NOT
#colnames and names are just the names of columns in a dataframe and a vector respectively
# "%in%" opterator means "in"
#We use the data[rows,columns] format to only select columns and save it as a new object
data.clean = df[,!colnames(df) %in% names(NAS.70)]


#Check to make sure there are no NAs
NAS.clean = sapply(data.clean, function(x) sum(is.na(x)))
sum(NAS.clean > 0)
```

Clean r environment
```{r}
rm(list= ls()[!(ls() %in% c("data.clean"))])
gc()
```

Convert to long format
```{r}
#First create a "long" data.cleanset with miRNA name in one column and expression in another
long.format = gather(data.clean, key = "miRNA", value = "expression", contains("mir"))

# convert to factors
long.format = long.format %>% mutate(Education = as.factor(Education),
                                     SES = as.factor(SES),
                                     Environmental_Smoking = as.factor(Environmental_Smoking),
                                     Child_Sex = as.factor(Child_Sex))
```

Model of mirna --> child cognition adjusted for covaraites
```{r}
#First run the model
fit.tidy = long.format %>% group_by(miRNA) %>% 
  do(tidy(lm(GCI ~ expression + Education + SES + Maternal_Age + Maternal_BMI + Environmental_Smoking + Child_Sex, data=.)))

#Then run the CIs
fit.tidy.CI = long.format %>% group_by(miRNA) %>% 
  do(as.data.frame(confint(lm(GCI ~ expression + Education + SES + Maternal_Age + Maternal_BMI + Environmental_Smoking + Child_Sex, data=.)))) %>%
  bind_cols(fit.tidy)

#FDR adjust p values
fit.tidy.CI$FDR.q = p.adjust(fit.tidy.CI$p.value, method = "fdr")

#Sum of sig FDR miRNAs
fit1 = fit.tidy.CI %>% 
  filter(term == "expression") 

#shows 36 significant mirnas
sum(fit1$FDR.q <= 0.05)

```

model of maternal violence --> mirna adjusted for covariates
```{r}
#First run the model
fit.tidy = long.format %>% group_by(miRNA) %>% 
  do(tidy(lm(expression ~ Rasch_ETV_Victim + Education + SES + Maternal_Age + Maternal_BMI + Environmental_Smoking + Child_Sex, data=.)))

#Then run the CIs
fit.tidy.CI = long.format %>% group_by(miRNA) %>% 
  do(as.data.frame(confint(lm(expression ~ Rasch_ETV_Victim + Education + SES + Maternal_Age + Maternal_BMI + Environmental_Smoking + Child_Sex, data=.)))) %>%
  bind_cols(fit.tidy)

#FDR adjust p values
fit.tidy.CI$FDR.q = p.adjust(fit.tidy.CI$p.value, method = "fdr")

#Sum of sig FDR miRNAs
fit2 = fit.tidy.CI %>% 
  filter(term == "Rasch_ETV_Victim") 

#shows 3 significant mirnas
sum(fit2$FDR.q <= 0.05)
```

model of materal violence --> child cognition adjusted for covariates
```{r}
fit = lm(GCI ~ Rasch_ETV_Witnessed + Education + SES + Maternal_Age + Maternal_BMI + Environmental_Smoking + Child_Sex, data = data.clean)

broomfit = broom::tidy(fit)
```



Formal mediation analysis: add the three miRNAs that significantly affected child cognition to model of maternal violence on child cognition. If this reduces the effect of maternal violence, then the miRNAs are mediators
```{r}
fitmed = lm(GCI ~ Rasch_ETV_Victim + Education + SES + Maternal_Age + Maternal_BMI + Environmental_Smoking + Child_Sex + mir_20a_A + mir_7_B + mir_942_B, data = data.clean)
```

Plot for figure 2
```{r}
df %>% ggplot(aes(y = GCI, x = Rasch_ETV_Victim)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(0,1) +
    labs(title = "Figure 2",
       x = "Rasch ETV Victim Score",
       y = "General Cognitive Index") +
  theme_bw() 

ggsave("figure 2.jpg",dpi=300)
```

